Appendix

### 2014-12-09 Rmarkdown/appendix.Rmd
Re-jigged appendix

## Equivalence of tests to identify DMCs in a one-group experiment

@@

## The probability that two dependent Bernoulli random variables are identical

@@

## Analysis of \cite{Lacey:2013iy}

I want to estimate the number of CpGs used by \cite{Lacey:2013iy} to estimate the correlation of methylation levels. They used the chromosome 11 data from an RRBS experiment on a normal myotube cell line, MTCTL2. Unfortunately, the data are not publicly available. Instead, I use data from ENCODE of a skeletal muscle myotubes cell line that had RRBS performed in triplicate (HSMMtube).

Firstly, I download the data and unzip it:
```{r, cache = FALSE, message = FALSE}
library(R.utils)
library(data.table)
library(dplyr)
print(getwd())
```

```{r, cache = FALSE, message = FALSE}
dl_dir <- 'downloads/'
url <- 'http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/'
files <- paste0('wgEncodeHaibMethylRrbsHsmmtubefshdDukeSitesRep',
                1:3, '.bed.gz')
dests <- paste0(dl_dir, files)
x <- mapply(function(f, d) {
  if (!file.exists(gsub(".gz", "", paste0(dl_dir, f)))) {
    download.file(url = paste0(url, f), d)
    gunzip(filename = d, overwrite = TRUE)
    }
  fread(gsub(pattern = '.gz', replacement = '', x = d), sep = '\t',
        verbose = FALSE)
  }, files, dests, SIMPLIFY = FALSE)
```

Then I count the number of CpGs on chromosome 11 in each sample:
```{r}
y <- lapply(x, function(xx) {
  xx <- tbl_dt(xx)
  xx %>%
    group_by(V1) %>%
    summarise(n())
})

rbind_all(y) %>% filter(V1 == 'chr11')
```

## Computing details

>>>>

\section{Equivalence of tests to identify DMCs in a one-group experiment}\label{app:one_group_equivalence}

\subsection{Mathematical equivalence}

@@

\subsection{Computational equiavelence}

\section{The probability that two dependent Bernoulli random variables are identical}\label{app:bernoulli_equal}

@@

\section{Estimate of number of CpGs used by \citet{Lacey:2013iy}}\label{app:lacey}

I want to estimate the number of CpGs used by \citet{Lacey:2013iy} to estimate the correlation of methylation levels. They used the chromosome 11 data from an RRBS experiment on a normal myotube cell line, _MTCTL2_. Unfortunately, there is no URL provided to these data. \citet{Lacey:2013iy} cite their data as having come from \citet{Tsumagari:2013gn}, which includes a sample "Normal Human Skeletal Muscle Myoblasts", _HSMM_ ([http://genome-mirror.duhs.duke.edu/ENCODE/cellTypes.html](http://genome-mirror.duhs.duke.edu/ENCODE/cellTypes.html), [http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/](http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/)). I believe this to be the same cell line. In any case, for my purposes I simply needed some comparable RRBS data, not necessarily the exact same sample as  _MTCTL2_.

There are approximately $60,000$ CpGs on chromosome 11 in these comparable RRBS data. The following R code shows how I computed this using three RRBS replicates of the _HSMM_ sample.

Firstly, I download the data and unzip it:
```{r, cache = FALSE, message = FALSE}
library(R.utils)
library(data.table)
dl_dir <- "../data/appendix/"
url <- 'http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeHaibMethylRrbs/'
files <- paste0('wgEncodeHaibMethylRrbsHsmmtubefshdDukeSitesRep',
                1:3, '.bed.gz')
dests <- paste0(dl_dir, files)
x <- mapply(function(f, d) {
  if (!file.exists(gsub(".gz", "", paste0(dl_dir, f)))) {
    download.file(url = paste0(url, f), d)
    gunzip(filename = d, overwrite = TRUE)
    }
  fread(gsub(pattern = '.gz', replacement = '', x = d), sep = '\t',
        verbose = FALSE)
  }, files, dests, SIMPLIFY = FALSE)
```

Then I count the number of CpGs on chromosome 11 in each replicate:
```{r}
y <- rbindlist(x)
y[V1 == "chr11", .N, by = list(V4)]
```

\section{Computing details}\label{app:computing_details}
